Bandit Multiclass Linear Classification: Efficient Algorithms for the Separable Case

### 2019-01-22 ICML-submission/algorithm-for-strongly-linearly-separable-data.tex
add one paragraph to intro + change algorithm style
Definition and reference added for positive definite kernels

\begin{algorithm}[h]
\caption{\textsc{Bandit Algorithm for Strongly Linearly Separable Examples}
\label{algorithm:algorithm-for-strongly-linearly-separable-examples}}
\begin{algorithmic}[1]
{
\REQUIRE{Number of classes $K$, number of rounds $T$.}
\REQUIRE{Inner product space $(V,\ip{\cdot}{\cdot})$.}
\STATE{Initialize $w_1^{(1)} = w_2^{(1)} = \dots = w_K^{(1)} = 0$}
\FOR{$t=1,2,\dots,T$}
\STATE{Observe feature vector $x_t \in V$}
\STATE{Compute $S_t = \left\{ i ~:~ 1 \le i \le K, \ \ip{w_i^{(t)}}{x_t} \ge 0 \right\}$}
\IF{$S_t = \emptyset$}
\STATE{Predict $\widehat y_t \sim \text{Uniform}(\{1,2,\dots,K\})$}
\STATE{Observe feedback $z_t = \indicator{\widehat y_t \neq y_t}$}
\IF{$z_t = 1$}
\STATE{Set $w_i^{(t+1)} = w_i^{(t)}$ for all $i \in \{1,2,\dots,K\}$}
\ELSE
\STATE{Set $w_i^{(t+1)} = w_i^{(t)}$ \\ \qquad  for all $i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}$}
\STATE{Update $w_{\widehat y_t}^{(t+1)} = w_{\widehat y_t}^{(t)} + x_t$}
\label{line:pos-update}
\ENDIF
\ELSE
\STATE{Predict $\widehat y_t \in S_t$ chosen arbitrarily}
\STATE{Observe feedback $z_t  = \indicator{\widehat y_t \neq y_t}$}
\IF{$z_t = 1$}
\STATE{Set $w_i^{(t+1)} = w_i^{(t)}$ \\ \qquad for all $i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}$}
\STATE{Update $w_{\widehat y_t}^{(t+1)} = w_{\widehat y_t}^{(t)} - x_t$}
\label{line:neg-update}
\ELSE
\STATE{Set $w_i^{(t+1)} = w_i^{(t)}$ for all $i \in \{1,2,\dots,K\}$}
\ENDIF
\ENDIF
\ENDFOR
}
\end{algorithmic}
\end{algorithm}

@@

Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples} can
be extended to nonlinear classification using
kernels~\cite{Scholkopf-Smola-2002, Shawe-Taylor-Cristianini-2004}. The formal
description of the kernelized algorithm is given in
Algorithm~\ref{algorithm:kernelized}. As opposed to explicitly maintaining the
weight vector for each class, the algorithm maintains the set of example-scalar
pairs corresponding to the updates of the non-kernelized algorithm. As a direct
consequence of
Theorem~\ref{theorem:strongly-separable-examples-mistake-upper-bound} we get a
mistake bound for the kernelized algorithm.

@@

\begin{algorithm}[h]
\caption{\textsc{Kernelized Bandit Algorithm}
\label{algorithm:kernelized}}
\begin{algorithmic}[1]
{
\REQUIRE{Number of classes $K$, number of rounds $T$.}
\REQUIRE{Kernel function  $k(\cdot, \cdot)$.}
\STATE{Initialize $J_1^{(1)} = J_2^{(1)} = \dots = J_K^{(1)} = \emptyset$}
\FOR{$t=1,2,\dots,T$}
\STATE{Observe feature vector $x_t$.}
\STATE{Compute \\ $S_t = \left\{ i ~:~ 1 \le i \le K, \ \sum_{(x,y) \in J_i^{(t)}} y k(x, x_t) \ge 0 \right\}$}
\IF{$S_t = \emptyset$}
\STATE{Predict $\widehat y_t \sim \text{Uniform}(\{1,2,\dots,K\})$}
\STATE{Observe feedback $z_t = \indicator{\widehat y_t \neq y_t}$}
\IF{$z_t = 1$}
\STATE{Set $J_i^{(t+1)} = J_i^{(t)}$ for all $i \in \{1,2,\dots,K\}$}
\ELSE
\STATE{Set $J_i^{(t+1)} = J_i^{(t)}$ \\ \qquad  for all $i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}$}
\STATE{Update $J_{\widehat y_t}^{(t+1)} = J_{\widehat y_t}^{(t)} \cup \cbr{(x_t, +1)}$}
\ENDIF
\ELSE
\STATE{Predict $\widehat y_t \in S_t$ chosen arbitrarily}
\STATE{Observe feedback $z_t =  \indicator{\widehat y_t \neq y_t}$}
\IF{$z_t = 1$}
\STATE{Set $J_i^{(t+1)} = J_i^{(t)}$ \\ \qquad for all $i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}$}
\STATE{Update $J_{\widehat y_t}^{(t+1)} = J_{\widehat y_t}^{(t)} \cup \cbr{(x_t, -1)}$}
\ELSE
\STATE{Set $J_i^{(t+1)} = J_i^{(t)}$ for all $i \in \{1,2,\dots,K\}$}
\ENDIF
\ENDIF
\ENDFOR
}
\end{algorithmic}
\end{algorithm}

>>>>

%\begin{algorithm}[h]
%\caption{\textsc{Bandit Algorithm for Strongly Linearly Separable Examples}
%\label{algorithm:algorithm-for-strongly-linearly-separable-examples}}
%\begin{algorithmic}[1]
%{
%\REQUIRE{Number of classes $K$, number of rounds $T$.}
%\REQUIRE{Inner product space $(V,\ip{\cdot}{\cdot})$.}
%\STATE{Initialize $w_1^{(1)} = w_2^{(1)} = \dots = w_K^{(1)} = 0$} \\
%\FOR{$t=1,2,\dots,T$}
%\STATE{Observe feature vector $x_t \in V$}
%\STATE{Compute $S_t = \left\{ i ~:~ 1 \le i \le K, \ \ip{w_i^{(t)}}{x_t} \ge 0 \right\}$}
%\IF{$S_t = \emptyset$}
%\STATE{Predict $\widehat y_t \sim \text{Uniform}(\{1,2,\dots,K\})$}
%\STATE{Observe feedback $z_t = \indicator{\widehat y_t \neq y_t}$}
%\IF{$z_t = 1$}
%\STATE{Set $w_i^{(t+1)} = w_i^{(t)}$ for all $i \in \{1,2,\dots,K\}$}
%\ELSE
%\STATE{Set $w_i^{(t+1)} = w_i^{(t)}$ \\ \qquad  for all $i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}$}
%\STATE{Update $w_{\widehat y_t}^{(t+1)} = w_{\widehat y_t}^{(t)} + x_t$}
%\label{line:pos-update}
%\ENDIF
%\ELSE
%\STATE{Predict $\widehat y_t \in S_t$ chosen arbitrarily}
%\STATE{Observe feedback $z_t  = \indicator{\widehat y_t \neq y_t}$}
%\IF{$z_t = 1$}
%\STATE{Set $w_i^{(t+1)} = w_i^{(t)}$ \\ \qquad for all $i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}$}
%\STATE{Update $w_{\widehat y_t}^{(t+1)} = w_{\widehat y_t}^{(t)} - x_t$}
%\label{line:neg-update}
%\ELSE
%\STATE{Set $w_i^{(t+1)} = w_i^{(t)}$ for all $i \in \{1,2,\dots,K\}$}
%\ENDIF
%\ENDIF
%\ENDFOR
%}
%\end{algorithmic}
%\end{algorithm}

\begin{algorithm}[h]
\SetAlgoLined
\LinesNumbered
\caption{\textsc{Bandit Algorithm for Strongly Linearly Separable Examples}
\label{algorithm:algorithm-for-strongly-linearly-separable-examples}}
\textbf{Require:} Number of classes $K$, number of rounds $T$. \\
\textbf{Require:} Inner product space $(V,\ip{\cdot}{\cdot})$.  \\
\nl Initialize $w_1^{(1)} = w_2^{(1)} = \dots = w_K^{(1)} = 0$\\ 
\nl \For{$t=1,2,\dots,T$}{
\nl    Observe feature vector $x_t \in V$ \\
\nl    Compute $S_t = \left\{ i ~:~ 1 \le i \le K, \ \ip{w_i^{(t)}}{x_t} \ge 0 \right\}$\\
\nl    \If{$S_t = \emptyset$}{
\nl         Predict $\widehat y_t \sim \text{Uniform}(\{1,2,\dots,K\})$ \\
\nl         Observe feedback $z_t = \indicator{\widehat y_t \neq y_t}$\\
\nl          \If{$z_t = 1$}{
\nl               Set $w_i^{(t+1)} = w_i^{(t)}$, $\forall i \in \{1,2,\dots,K\}$
              }
\nl          \Else{
\nl               Set $w_i^{(t+1)} = w_i^{(t)}$, $\forall i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}$ \\
\nl Update $w_{\widehat y_t}^{(t+1)} = w_{\widehat y_t}^{(t)} + x_t$  \label{line:pos-update}      
              }
        }
\nl     \Else{
\nl           Predict $\widehat y_t \in S_t$ chosen arbitrarily \\
\nl           Observe feedback $z_t  = \indicator{\widehat y_t \neq y_t}$ \\
\nl            \If{$z_t = 1$}{
\nl                Set $w_i^{(t+1)} = w_i^{(t)}$, $\forall i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}$ \\
\nl  Update $w_{\widehat y_t}^{(t+1)} = w_{\widehat y_t}^{(t)} - x_t$   \label{line:neg-update}            
               }
\nl           \Else{
\nl                 Set $w_i^{(t+1)} = w_i^{(t)}$, $\forall i \in \{1,2,\dots,K\}$
           }
      }
}

\end{algorithm}

@@

Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples} can
be extended to nonlinear classification using
\emph{positive definite kernels} (or \emph{kernels}, for short), which are 
functions of the form $k: X^2 \to V$ for some set $X$ and inner product space 
$(V, \ip{\cdot}{\cdot})$ such that the matrix $\left(k(x_i,x_j)\right)_{i,j=1}^m$ 
is positive semidefinite for any positive integer $m$ and $x_1, \dots, x_m \in X$
~\citep[][Definition 2.5]{Scholkopf-Smola-2002}. 
%kernels~\cite{Scholkopf-Smola-2002, Shawe-Taylor-Cristianini-2004}. The formal
%description of the kernelized algorithm is given in
%Algorithm~\ref{algorithm:kernelized}. 
As opposed to explicitly maintaining the
weight vector for each class, the algorithm maintains the set of example-scalar
pairs corresponding to the updates of the non-kernelized algorithm. As a direct
consequence of
Theorem~\ref{theorem:strongly-separable-examples-mistake-upper-bound} we get a
mistake bound for the kernelized algorithm.

@@

\begin{algorithm}[h]
\caption{\textsc{Kernelized Bandit Algorithm}
\label{algorithm:kernelized}}
\textbf{Require:} Number of classes $K$, number of rounds $T$.\\
\textbf{Require:} Kernel function  $k(\cdot, \cdot)$. \\
Initialize $J_1^{(1)} = J_2^{(1)} = \dots = J_K^{(1)} = \emptyset$ \\
\For{$t=1,2,\dots,T$} {
     Observe feature vector $x_t$.\\
     Compute \\ $S_t = \left\{ i ~:~ 1 \le i \le K, \ \sum_{(x,y) \in J_i^{(t)}} y k(x, x_t) \ge 0 \right\}$ \\
     \If{$S_t = \emptyset$}{
           Predict $\widehat y_t \sim \text{Uniform}(\{1,2,\dots,K\})$ \\
           Observe feedback $z_t = \indicator{\widehat y_t \neq y_t}$ \\
           \If{$z_t = 1$}{
                Set $J_i^{(t+1)} = J_i^{(t)}$ for all $i \in \{1,2,\dots,K\}$
           }
           \Else{
                Set $J_i^{(t+1)} = J_i^{(t)}$, $\forall i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}$ \\
                Update $J_{\widehat y_t}^{(t+1)} = J_{\widehat y_t}^{(t)} \cup \cbr{(x_t, +1)}$
           }
      }
      \Else{
           Predict $\widehat y_t \in S_t$ chosen arbitrarily \\
           Observe feedback $z_t =  \indicator{\widehat y_t \neq y_t}$ \\
           \If{$z_t = 1$}{
                Set $J_i^{(t+1)} = J_i^{(t)}$, $\forall i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}$\\
                Update $J_{\widehat y_t}^{(t+1)} = J_{\widehat y_t}^{(t)} \cup \cbr{(x_t, -1)}$ 
           }
           \Else {
                Set $J_i^{(t+1)} = J_i^{(t)}$ for all $i \in \{1,2,\dots,K\}$
           }
      }
}

\end{algorithm}
